---
title: "Statistical Simulation"
author: "Leihua Ye"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# statistical simulation 101
# part i: basics of sampling
- runif 
- sample
- sapply
- apply
- applications

Check the sample function; it is the foundation of statistical modeling 

sample(x, size, replace = FALSE, prob = NULL)

x: a vector of one or more elements from which to choose, or a positive integer
size: a non-negative integer giving the number of items to choose
replace: should sampling be with replacement?
prob: a vector of probability weights for obtaining the elements of the vector being sampled 

# how to generate a random sample
runif(n,min=minimal_value,max=max_value)
```{r}
set.seed(2)
runif(25,min=0,max=10)

runif(25,min=0,max=10) %>% 
  round(.,digits = 0)
```


```{r}
# flip a coin 
sample(c('H','T'),size = 10,replace=TRUE)
sample(c(1:6),size=10,replace=TRUE)
```

# the sample() function draws randomly from a specified set of (scalar) objects allowing you to sample from arbitrary distributions of numbers
```{r}
#1. random permutation of sequence [1,10]
set.seed(2)
sample(10)
?sample()
```

```{r}
sample(1:10,4)
```

```{r}
set.seed(1)
sample(letters,18)
```

```{r}
# from r documentation
x<- 1:10 
#a random permutation
sample(x)

# resample with replacement
sample(x,replace=TRUE)
```

```{r}
# random sample of size 10 from sequence [1,5] with equal probabilities
equal_prob_dist = sample(5,10000,prob=rep(0.1,5),replace=T)
hist(equal_prob_dist)
```

```{r}
#random sample of size 10 from sequence[1,5] with unequal probabilities 
unequal_prob_dist = sample(5,10000,prob = c(0.1,0.25,0.4,0.25,0.1),replace=T)
hist(unequal_prob_dist)
```
By default, the probability is equal if don't specify it.

# To sample rows from a data frame or a list, we can sample the indices into an object rather than the elements of the object itself. 取元素对应的indices！
```{r}
head(mtcars)
```
```{r}
# create an index vector for the elements/rows 
index <- seq_len(nrow(mtcars))
 
#sample from the index vector 
set.seed(12) 

#to obtain a random sample of 10
sample_index <- sample(index,10)

# to show the sampled elements/rows 
mtcars[sample_index,]
```



# Part 2: Application
# example 1: dies
```{r}
# use sample() to run 10,000 trials using two fair dies. what is the probability of rolling a 7?
set.seed(1)
die = 1:6
die1 = sample(die,10000,replace = TRUE,prob=NULL)
die2= sample(die,10000,replace=TRUE,prob = NULL)
outcomes = die1+die2
mean(outcomes == 7)
```

```{r}
# simulate the process 10000 times and check for discrepency 
set.seed(1)
for (i in 10000){
  die_1 = sample(die,prob=NULL,replace=TRUE)
  die_2 = sample(die,prob=NULL,replace=TRUE)
  die_sum = die_1+die_2
  print(mean(die_sum==7))
}
```
```{r}
# check for system time
system.time(for (i in 10000){
  die_1 = sample(die,prob=NULL,replace=TRUE)
  die_2 = sample(die,prob=NULL,replace=TRUE)
  die_sum = die_1+die_2
  print(mean(die_sum==7))
})
```

```{r}
#You have two dies, and what is the probability of rolling a 7?
# use sample() to run 10,000 trials using two fair dies. 
set.seed(1)
die = 1:6
die_1 = sample(die,20000,replace=TRUE,prob=rep(1/6,6))
die_2 = sample(die,20000,replace=TRUE,prob=rep(1/6,6))
outcomes = die_1+die_2
mean(outcomes==7)
```

```{r}
# What is the probability of rolling a 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, and 13?
sapply(2:13,function(x) mean(outcomes==x))
#?sapply()
```
# check for sapply()
sapply is a user-friendly version and wrapper of lapply by default returning a vector, matrix or, if simplify = "array", an array if appropriate, by applying simplify2array(). sapply(x, f, simplify = FALSE, USE.NAMES = FALSE) is the same as lapply(x, f).


```{r}
# Use the sample() function to determine the probability of rolling a 7 using three fair six-sided dies.
set.seed(1)
die = 1:6
die1 = sample(die,10000,replace=TRUE, prob = NULL)
die2 = sample(die1,10000,replace=TRUE, prob = NULL)
die3 = sample(die,10000,replace = TRUE, prob= NULL)
outcomes2 = die1+die2+die3
mean(outcomes2==7)
```

```{r}
# Use the sample() function to determine the probability of rolling a 7 using three fair six-sided dies.
set.seed(1)
die=1:6
die_1 = sample(die, 20000,replace=TRUE,prob=NULL)# by default, equal probability
die_2 = sample(die,20000,replace=TRUE, prob=NULL)
die_3 = sample(die,20000,replace=TRUE, prob=NULL)
die_combn = die_1+die_2+die_3
mean(die_combn==7)
```

```{r}
# Using three dies, what is the probability of rolling a 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, and 13?
sapply(2:13,function(x) mean(die_combn==x))
```

# example 2 
```{r}
# do 10k samples and calculate how many observations are included and how many are not included? 
set.seed(1)
n=10000

included_obs = length(unique(sample(1:n, replace = TRUE)))
included_obs

missing_obs = n-included_obs
missing_obs
```

```{r}
# from 1:100, do a 10k sample and calculate how many observations are included and how many are not included? 
set.seed(1)
n = 100
included_observations = length(unique(sample(1:100,replace=TRUE, prob=NULL)))
included_observations/n

(n-included_observations)/n
```

#3 example 3
```{r}
# three ways of generating a m*n matrix with randomly assigned 0/1

#3.1 for loop
# create an empty matrix
m <- 10
n <- 10
m00 <- matrix(0,m,n)

for (i in 1:m) {
  for (j in 1:n) {
    m00[i,j] <-sample(c(0,1),1)
  }
}

m00
```

```{r}
system.time(for (i in 1:m) {
  for (j in 1:n) {
    m00[m,n] <-sample(c(0,1),1)
  }
}
)
```



```{r}
#3.2 apply() function
m <-10
n<-10

m0 <- matrix(0,m,n)

apply(m0,c(1,2),function(x) sample(c(0,1),1))

system.time(apply(m0,c(1,2),function(x) sample(c(0,1),1)))
?apply()
```
apply():
- Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.
- apply(x,margin, fun, ...)
    - x: an array, or matrix 
    - margin: a vector giving the subscripts which the function will be applied over. for a matrix 1 indicates rows, and 2 indicates columns. c(1,2) indicates rows and columns. where x has named dimnames, it can be a character vector selecting dimension names 
    - fun: the function to be applied. 

```{r}
#3.3 some other methods 
#1 generate a bunch of uniformly distributed[0,1) random numbers; round them to the closest integer
m1<-round(matrix(runif(r*c),r,c))
#m1
system.time(m1<-round(matrix(runif(r*c),r,c)))
```

```{r}
#generate r*c random numbers following a binomial distribution; 
# allow for different probabilities, rahter than 0.5 as in m1
m2 <- matrix(rbinom(r*c,1,0.5),r,c)
#m2
```

```{r}
system.time(m3<-matrix(round(runif(r*c)),r,c))
```

```{r}
m4<-matrix(sample(0:1,r*c,replace=TRUE),r,c)
#m4
```

#4 example 4
Flip a coin 10 times and simulate the process for 10,000 times. Show the distribution of the number of heads shown up.
```{r}
# create an empty list 
total_heads = c()

# use a for looop to simulate coin-flipping 10 times; repeat it for 10,000 times. 
for (i in 1:10000){
  sum_heads = sum(round(runif(10,0,1)))
  total_heads = c(total_heads, sum_heads)
  }

hist(total_heads)
```

Reference:
1. https://bookdown.org/rdpeng/rprogdatascience/simulation.html#random-sampling
2. R documentation of Sampling 
3. https://crumplab.github.io/programmingforpsych/simulating-and-analyzing-data-in-r.html
